---
title: "Homework_problem_2"
author: "May Dixon"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, echo=FALSE}
R<- read.csv("https://raw.githubusercontent.com/difiore/ADA2016/master/zombies.csv", sep=",", header=TRUE)
```
##1.  
 Calculate the population mean and standard deviation for each quantitative random variable (height, weight, age, number of zombies killed, and years of education). NOTE: You will not want to use the built in var() and sd() commands as these are for samples.
 
 
```{r, echo=TRUE}
attach(R)
mean_z<- function(x) { #generic mean function
      sum(x)/length(x) 
}
sd_z<-function(x){ #population standard deviation function
      sqrt(sum((x-mean_z(x))^2)/length(x))
}
```

```{r, include = FALSE, echo=FALSE}
library(dplyr)
```
Means and standard deviations of height, weight, zombies killed, years of education, and age:
```{r, echo=FALSE}
summarise(R, mean_z(height) , sd_z(height),mean_z(weight), sd_z(weight), mean_z(zombies_killed), sd_z(zombies_killed), mean_z(years_of_education), sd_z(years_of_education), mean_z(age), sd_z(age) )
```

```{r, include=FALSE, eval=FALSE}
#took this out bc more efficient way to do it
# <!-- Mean and SD of height:    -->
# <!-- ```{r} -->
# <!-- # Mean and standard deviation of height -->
# <!-- mean_z(height) -->
# <!-- sd_z(height) -->
# <!-- ``` -->
# <!-- Mean and SD of weight:    -->
# <!-- ```{r} -->
# <!-- #mean and sd of weight -->
# <!-- mean_z(weight) -->
# <!-- sd_z(weight) -->
# <!-- ``` -->
# <!-- Mean and SD of zombies killed:    -->
# <!-- ```{r} -->
# <!-- #mean and sd of zombies killed -->
# <!-- mean_z(zombies_killed) -->
# <!-- sd_z(zombies_killed) -->
# <!-- ``` -->
# <!-- Mean and SD of education years:    -->
# <!-- ```{r} -->
# <!-- # mean and sd of years of education -->
# <!-- mean_z(years_of_education) -->
# <!-- sd_z(years_of_education) -->
# <!-- ``` -->
# <!-- Mean and SD of age:    -->
# <!-- ```{r} -->
# <!-- # mean and sd of age -->
# <!-- mean_z(age) -->
# <!-- sd_z(age) -->
# <!-- ``` -->
```
##2. 
Make boxlots with ggplot by gender:

```{r, echo=FALSE}
library(ggplot2)
hg <- ggplot(data = R, aes(x = gender, y = height))
hg <- hg + geom_boxplot()
hg <- hg + theme(axis.text.x = element_text(angle = 90))
hg <- hg + ylab("Height (inches)")
hg

wg <- ggplot(data = R, aes(x = gender, y = weight))
wg <- wg + geom_boxplot()
wg <- wg + theme(axis.text.x = element_text(angle = 90))
wg <- wg + ylab("Weight (lbs)")
wg

zg <- ggplot(data = R, aes(x = gender, y = zombies_killed))
zg <- zg + geom_boxplot()
zg <- zg + theme(axis.text.x = element_text(angle = 90))
zg <- zg + ylab("Number of zombies killed")
zg

eg <- ggplot(data = R, aes(x = gender, y = years_of_education))
eg <- eg + geom_boxplot()
eg <- eg + theme(axis.text.x = element_text(angle = 90))
eg <- eg + ylab("Years of education")
eg

ag <- ggplot(data = R, aes(x = gender, y = age))
ag <- ag + geom_boxplot()
ag <- ag + theme(axis.text.x = element_text(angle = 90))
ag <- ag + ylab("Age of survivors")
ag
```
  
  Hooray egalitarian universe!
  
##3.  
Use ggplot and make a scatterplots of height and weight in relation to age. Do these variables seem to be related? In what way?
```{r, echo=FALSE}
ah <- ggplot(data = R, aes(x = age, y = height))
ah <- ah + geom_point()
ah <- ah + xlab("Height of survivors")
ah <- ah + ylab("age of survivors")
ah <- ah + geom_smooth(method=lm)
ah

aw <- ggplot(data = R, aes(x = age, y = weight))
aw <- aw + geom_point()
aw <- aw + xlab("Age of survivors")
aw <- aw + ylab("Weight of survivors")
aw <- aw + geom_smooth(method=lm)
aw
```
    
Both height and weight seem to be positively correlated with age, though the correlation seems to be stronger with height than age   

##4. 
Using histograms and Q-Q plots, check whether the quantitative variables seem to be drawn from a normal distribution. Which seem to be and which do not (hint: not all are drawn from the normal distribution)? For those that are not, can you determine what common distribution they are drawn from?

```{r, echo=FALSE}
hist(height)
qqnorm(height)

hist(weight)
qqnorm(weight)

hist(zombies_killed, xlim=range(0,12))
qqnorm(zombies_killed)

hist(years_of_education)
qqnorm(years_of_education)

hist(age)
qqnorm(age)

```

Height, weight, and age seem to be drawn from  normal distributions, whereas zombies killed and years of education do not. These variables are both right skewed, and follow **poisson distributions**. 

##5.
Now use the sample() function to sample ONE subset of 30 zombies (without replacement) from this population and calculate the mean and sample standard deviation for each variable. Also estimate the standard error for each variable and construct the 95% confidence interval for each mean. Note that for the variables that are not drawn from the normal distribution, you will need to base your estimate of the CIs on some different distribution.

```{r, echo=FALSE}

Z_sam <- R[sample(1:1000, 30,
  	replace=FALSE),] #sampling 30 rows of the greater population
detach(R)
attach(Z_sam)
```

####Mean, standard deviation, standard error, and 95% confidence interval for height:

```{r}
mzs <- mean(height) #mean 
sdzs <- sd(height) #standard deviation
sdzs
sqrt(var(height)) #standard deviation 2
sqrt(sum((height-mean_z(height))^2)/(length(height)-1)) #standard deviation 3
#all of these have the same output/ use the same variable, so I'm just going to use the built in one, sqrt
m_pop<-mean(R$height)
m_pop

se_z<-function(x){ #function for determining standard error of the means, just define m_pop outisde function
    m_sam<-mean(x)
    sqrt(((m_sam - m_pop)^2)/1)
    
}
se_z(height)

#constructing 95% confidence interval
n<-30
error <- qnorm(0.975)*sdzs/sqrt(n)
left <- mzs-error
right <- mzs+error
left 
right
```

There is a 95% chance that the true mean falls between 68.13 and 71.19

#### Mean, standard deviation, standard error, and 95% confidence interval for weight:
```{r}
mzs <- mean(weight) #mean
mzs
sdzs <- sd(weight) #standard deviation
sdzs

m_pop<-mean(R$weight)

se_z(weight) #standard error of the mean

#constructing 95% confidence interval
n<-30
error <- qnorm(0.975)*sdzs/sqrt(n)
left <- mzs-error
right <- mzs+error
left #left tail
right #right tail
```
  
  There is a 95% chance that a value frawn from this distribution will lie between 143.93 and 159.54
  
#### Mean, standard deviation, standard error, and 95% confidence interval for zombies killed:
```{r}
mzs <- mean(zombies_killed) #mean
mzs
sdzs <- sd(zombies_killed) #standard deviation
sdzs

m_pop<-mean(R$zombies_killed)

se_z(zombies_killed) #standard error of the mean

#constructing 95% confidence interval
n<-30
lambdaz <- (sum(zombies_killed))/30 #lambda for this dist (total zombies killed/ # people)
error <- qpois(0.975, lambda=lambdaz)*sdzs/sqrt(n)
left <- mzs-error
right <- mzs+error
left #left tail
right #right tail
```
  
  There is a 95% chance that the true mean falls between  0.5 and 4.3
  
#### Mean, standard deviation, standard error, and 95% confidence interval for years of education:
```{r}
mzs <- mean(years_of_education) #mean
mzs
sdzs <- sd(years_of_education) #standard deviation
sdzs

m_pop<-mean(R$years_of_education)

se_z(years_of_education) #standard error of the mean

#constructing 95% confidence interval
n<-30
lambdaz <- (sum(years_of_education))/30 #lambda for this dist (total years education / # people)
error <- qpois(0.975, lambda=lambdaz)*sdzs/sqrt(n)
left <- mzs-error
right <- mzs+error
left #left tail
right #right tail
```
  
  There is a 95% chance that the true mean falls between 0.8 and 4.9
  
#### Mean, standard deviation, standard error, and 95% confidence interval for age:
```{r}
mzs <- mean(age) #mean
mzs
sdzs <- sd(age) #standard deviation of sample
sdzs

m_pop<-mean(R$age)

se_z(age) #standard error of the mean

#constructing 95% confidence interval
n<-30
error <- qnorm(0.975)*sdzs/sqrt(n)
left <- mzs-error
right <- mzs+error
left #left tail
right #right tail
```
  
  There is a 95% chance that the true mean falls between 19.7 and 21.7
  
##6.
Now draw 99 more random samples of 30 zombies out and calculate the mean for each of the these samples. Together with the first sample you drew out, you now have a set of 100 means for each variable (each based on 30 observations), which constitutes a sampling distribution for each variable.



```{r}
detach(Z_sam)
library(dplyr)
#########
Z_sam_list = list()
mean_heightlist <- NULL
mean_weightlist<- NULL
mean_zombiekilllist<- NULL
mean_educationlist<- NULL
mean_agelist<- NULL

for (i in 1:99) {
    # ... make some data
    z_sam_i <- R[sample(1:1000, 30,replace=FALSE),]
    #Z_sam_list$i <- i  # maybe you want to keep track of which iteration produced it? maybe doesn't matter to save
    #Z_sam_list[[i]] <- z_sam_i # add to list  #again doesn't matter
    mean_heightlist[i]<-mean(z_sam_i$height)
    mean_weightlist[i]<-mean(z_sam_i$weight)
    mean_zombiekilllist[i]<-mean(z_sam_i$zombies_killed)
    mean_educationlist[i]<-mean(z_sam_i$years_of_education)
    mean_agelist[i]<-mean(z_sam_i$age)
    
}
##########
zombie_means<- cbind(mean_heightlist, mean_weightlist, mean_zombiekilllist,mean_educationlist, mean_agelist)
#add the sample we already have
samp1<- c(69.66241, 151.7363, 2.366667, 2.8, 20.70173)
zombie_means <- rbind(zombie_means, samp1)

```

What are the means and standard deviations of this distribution for each variable? How do the standard deviations compare to the standard errors estimated in [5]? What do these sampling distributions look like? Are they normally distributed? What about for those variables that you concluded were not originally drawn from a normal distribution?

Means and st. devs of distribution of means:

```{r, echo=FALSE}
library(dplyr)

zombie_means<-as.data.frame(zombie_means) #this is th means of the bootstrapping efforts

s <- summarise(zombie_means, avgH = mean(mean_heightlist), sdH =sd(mean_heightlist),  avgW = mean(mean_weightlist), sdW = sd(mean_weightlist), avgZ = mean(mean_zombiekilllist), sdZ=sd(mean_zombiekilllist), avgE = mean(mean_educationlist), sdE = sd(mean_educationlist), avgA = mean(mean_agelist), sdA=sd(mean_agelist))
s
```

Can compare to the summary stats for #5:
```{r, echo=FALSE}
#can compare to the summary for #5 alone:
s_1_1samp<- summarise(Z_sam, avgH=mean(height), sdH=sd(height), avgW=mean(weight), sdW=sd(weight), avgZ=mean(zombies_killed), sdZ=sd(zombies_killed), avgE=mean(years_of_education), sdE=sd(years_of_education), avgA=mean(age), sdA=sd(age) )
s_1_1samp
```

And also to the true population means:
```{r, echo=FALSE}
#and to the true population values
s_1_pop<- summarise(R, avgH=mean(height), avgW=mean(weight),  avgZ=mean(zombies_killed), avgE=mean(years_of_education),  avgA=mean(age) )
s_1_pop
```

^ The bootstrapped means are closer to the population means, and have much smaller standard deviations.

v   The distribution of means drawn from other distributions (zombies killed) look normally distributed now. 
```{r, echo=FALSE}
#hists for distribution of means 
hist(mean_weightlist)
hist(mean_heightlist)
hist(mean_zombiekilllist)
hist(mean_educationlist)
hist(mean_agelist)
```

#                                    ---- El FIN ----
