---
title: "Homework_week_4"
author: "May Dixon"
date: "September 28, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
### Prompt  

Every Saturday, at the same time, a primatologist goes and sits in the forest in the morning and listens for titi monkey calls, counting the number of calls they hear in a 2 hour window from 5am to 7am. Based on previous knowledge, she believes that the mean number calls she will hear in that time is exactly 15. Let X represent the appropriate Poisson random variable of the number of calls heard in each monitoring session.

```{r}

X<- rpois(1000, lambda = 15)

hist(X)

```


#### What is the probability that she will hear more than 8 calls during any given session?

```{r}
1-ppois(8, lambda = 15)

```

There is a 96.2% probability that the number of calls will be more than 8.   
  
#### What is the probability that she will hear no calls in a session?

```{r}
dpois(x=0,lambda=15)

```
There is a 3.06E-05% chance that she will hear no calls in a session.     

####What is the probability that she will hear exactly 3 calls in a session?

```{r}
dpois(x=3, lambda = 15)
```
There is a .017% chance she will hear exactly 3 calls.    

####Plot the relevant Poisson mass function over the values in range 0 ≤ x ≤ 30.
```{r}
#graph mass function
barplot(dpois(x=0:30,lambda=15),ylim=c(0,0.15),space=0, names.arg=0:30, ylab="Pr(X=x)", xlab="Number calls heard in 2 hours", main= "Poisson mass function of titi calls heard in 2 hours")

```

####Simulate 104 results from this distribution (i.e., 2 years of Saturday monitoring sessions).
```{r}

Sim.call<- rpois(104, lambda = 15)

```

####Plot the simulated results using hist() and use xlim() to set the horizontal limits to be from 0 to 30. How does your histogram compare to the shape of the probability mass function you plotted above?
```{r}
hist(Sim.call, xlim = c(0,30), xlab="Number of calls", main = "Freq. of hearing titi calls in 2 years of Saturdays")  
```
  
  The histogram is a lot less smooth  than the probability mass function above. (Life is random!)  
